\cchapter{2920, Last Year of the First Era}{Carlovac Townway}
\csection[Book]{1: Morning Star}
\cnote{1 Morning Star, 2920: Mournhold, Morrowind}
\dropcap{A}{lmalexia} lay in her bed of fur, dreaming. Not until the sun burned through her window, infusing the light wood and flesh colors of her chamber in a milky glow did she open her eyes. It was quiet and serene, a stunning reverse of the flavor of her dreams, so full of blood and celebration. For a few moments, she simply stared at the ceiling, trying to sort through her visions.

In the courtyard of her palace was a boiling pool which steamed in the coolness of the winter morning. At the wave of her hand, it cleared and she saw the face and form of her lover Vivec in his study to the north. She did not want to speak right away: he looked so handsome in his dark red robes, writing his poetry as he did every morning.

"Vivec," she said, and he raised his head in a smile, looking at her face across thousands of miles. "I have seen a vision of the end of the war."

"After eighty years, I don't think anyone can imagine an end," said Vivec with a smile, but he grew serious, trusting Almalexia's prophecies. "Who will win? Morrowind or the Cyrodilic Empire?"

"Without Sotha Sil in Morrowind, we will lose," she replied.

"My intelligence tells me the Empire will strike us to the north in early springtide, by First Seed at the latest. Could you go to Artaeum and convince him to return?"

"I'll leave today," she said, simply.

\cnote{4 Morning Star, 2920: Gideon, Black Marsh}
The Empress paced around her cell. Wintertide gave her wasteful energy, while in the summer she would merely sit by her window and be grateful for each breath of stale swamp wind that came to cool her. Across the room, her unfinished tapestry of a dance at the Imperial Court seemed to mock her. She ripped it from its frame, tearing the pieces apart as they drifted to the floor.

Then she laughed at her own useless gesture of defiance. She would have plenty of time to repair it and craft a hundred more. The Emperor had locked her up in Castle Giovesse seven years ago, and would likely keep her here until he or she died.

With a sigh, she pulled the cord to call her knight, Zuuk. He appeared at the door within minutes, fully uniformed as befitted an Imperial Guard. Most of the native Kothringi tribesmen of Black Marsh preferred to go about naked, but Zuuk had taken a positive delight to fashion. His silver, reflective skin was scarcely visible, only on his face, neck, and hands.

"Your Imperial Highness," he said with a bow.

"Zuuk," said Empress Tavia. "I'm bored. Lets discuss methods of assassinating my husband today."

\cnote{14 Morning Star, 2920: The Imperial City, Cyrodiil}
The chimes proclaiming South Wind's Prayer echoed through the wide boulevards and gardens of the Imperial City, calling all to their temples. The Emperor Reman III always attended a service at the Temple of the One, while his son and heir Prince Juilek found it more political to attend a service at a different temple for each religious holiday. This year, it was at the cathedral Benevolence of Mara.

The Benevolence's services were mercifully short, but it was not until well after noon that the Emperor was able to return to the palace. By then, the arena combatants were impatiently waiting for the start of the ceremony. The crowd was far less restless, as the Potentate Versidue-Shaie had arranged for a demonstration from a troupe of Khajiiti acrobats.

"Your religion is so much more convenient than mine," said the Emperor to his Potentate by way of an apology. "What is the first game?"

"A one-on-one battle between two able warriors," said the Potentate, his scaly skin catching the sun as he rose. "Armed befitting their culture."

"Sounds good," said the Emperor and clapped his hands. "Let the sport commence!"

As soon as he saw the two warriors enter the arena to the roar of the crowd, Emperor Reman III remembered that he had agreed to this several months before and forgotten about it. One combatant was the Potentate's son, Savirien-Chorak, a glistening ivory-yellow eel, gripping his katana and wakizashi with his thin, deceptively weak looking arms. The other was the Emperor's son, Prince Juilek, in ebony armor with a savage Orcish helm, shield and longsword at his side.

"This will be fascinating to watch," hissed the Potentate, a wide grin across his narrow face. "I don't know if I've even seen a Cyrodiil fight an Akavir like this. Usually it's army against army. At last we can settle which philosophy is better -- to create armor to combat swords as your people do, or to create swords to combat armor as mine do."

No one in the crowd, aside from a few scattered Akaviri counselors and the Potentate himself wanted Savirien-Chorak to win, but there was a collective intake of breath at the sight of his graceful movements. His swords seemed to be a part of him, a tail coming from his arms to match the one behind him. It was a trick of counterbalance, allowing the young serpent man to roll up into a circle and spin into the center of the ring in offensive position. The Prince had to plod forward the less impressive traditional way.

As they sprang at each other, the crowd bellowed with delight. The Akaviri was like a moon in orbit around the Prince, effortlessly springing over his shoulder to attempt a blow from behind, but the Prince whirled around quickly to block with his shield. His counter-strike met only air as his foe fell flat to the ground and slithered between his legs, tripping him. The Prince fell to the ground with a resounding crash.

Metal and air melted together as Savirien-Chorak rained strike after strike upon the Prince, who blocked every one with his shield.

"We don't have shields in our culture," murmured Versidue-Shaie to the Emperor. "It seems strange to my boy, I imagine. In our country, if you don't want to get hit, you move out of the way."

When Savirien-Chorak was rearing back to begin another series of blinding attacks, the Prince kicked at his tail, sending him falling back momentarily. In an instant, he had rebounded, but the Prince was also back on his feet. The two circled one another, until the snake man spun forward, katana extended. The Prince saw his foe's plan, and blocked the katana with his longsword and the wakizashi with his shield. Its short punching blade impaled itself in the metal, and Savirien-Chorak was thrown off balance.

The Prince's longblade slashed across the Akavir's chest and the sudden, intense pain caused him to drop both his weapons. It a moment, it was over. Savirien-Chorak was prostate in the dust with the Prince's longsword at his throat.

"The game's over!" shouted the Emperor, barely heard over the applause from the stadium.

The Prince grinned and helped Savirien-Chorak up and over to a healer. The Emperor clapped his Potentate on the back, feeling relieved. He had not realized when the fight had begun how little chance he had given his son at victory.

"He will make a fine warrior," said Versidue-Shaie. "And a great emperor."

"Just remember," laughed the Emperor. "You Akaviri have a lot of showy moves, but if just one of our strikes comes through, it's all over for you."

"Oh, I'll remember that," nodded the Potentate.

Reman thought about that comment for the rest of the games, and had trouble fully enjoying himself. Could the Potentate be another enemy, just as the Empress had turned out to be? The matter would bear watching.

\cnote{21 Morning Star, 2920: Mournhold, Morrowind}
"Why don't you wear that green gown I gave you?" asked the Duke of Mournhold, watching the young maiden put on her clothes.

"It doesn't fit," smiled Turala. "And you know I like red."

"It doesn't fit because you're getting fat," laughed the Duke, pulling her down on the bed, kissing her breasts and the pouch of her stomach. She laughed at the tickles, but pulled herself up, wrapping her red robe around her.

"I'm round like a woman should be," said Turala. "Will I see you tomorrow?"

"No," said the Duke. "I must entertain Vivec tomorrow, and the next day the Duke of Ebonheart is coming. Do you know, I never really appreciated Almalexia and her political skills until she left?"

"It is the same with me," smiled Turala. "You will only appreciate me when I'm gone."

"That's not true at all," snorted the Duke. "I appreciate you now."

Turala allowed the Duke one last kiss before she was out the door. She kept thinking about what he said. Would he appreciate her more or less when he knew that she was getting fat because she was carrying his child? Would he appreciate her enough to marry her?

\csection[Book]{2: Sun's Dawn}
\cnote{3 Sun's Dawn, 2920: The Isle of Artaeum, Summurset}
Sotha Sil watched the initiates float one by one up to the oassom tree, taking a fruit or a flower from its high branches before dropping back to the ground with varying degrees of grace. He took a moment while nodding his head in approval to admire the day. The whitewashed statue of Syrabane, which the great mage was said to have posed for in ancient days, stood at the precipice of the cliff overlooking the bay. Pale purple proscato flowers waves to and fro in the gentle breeze. Beyond, ocean, and the misty border between Artaeum and the main island of Summurset.

"By and large, acceptable," he proclaimed as the last student dropped her fruit in his hand. With a wave of his hand, the fruit and flowers were back in the tree. With another wave, the students had formed into position in a semicircle around the sorcerer. He pulled a small fibrous ball, about a foot in diameter from his white robes.

"What is this?"

The students understood this test. It asked them to cast a spell of identification on the mysterious object. Each initiate closed his or her eyes and imagined the ball in the realm of the universal Truth. Its energy had a unique resonance as all physical and spiritual matter does, a negative aspect, a duplicate version, relative paths, true meaning, a song in the cosmos, a texture in the fabric of space, a facet of being that has always existed and always will exist.

"A ball," said a young Nord named Welleg, which brought giggles from some of the younger initiates, but a frown from most, including Sotha Sil.

"If you must be stupid, at least be amusing," growled the sorcerer, and then looked at a young, dark-haired Altmer lass who looked confused. "Lilatha, do you know?"

"It's grom," said Lilatha, uncertainly. "What the dreugh meff after they've k-k-kr-krevinasim."

"Karvinasim, but very good, nonetheless," said Sotha Sil. "Now, tell me, what does that mean?"

"I don't know," admitted Lilatha. The rest of the students also shook their heads.

"There are layers to understanding all things," said Sotha Sil. "The common man looks at an object and fits it into a place in his way of thinking. Those skilled in the Old Ways, in the way of the Psijic, in Mysticism, can see an object and identify it by its proper role. But one more layer is needed to be peeled back to achieve understanding. You must identify the object by its role and its truth and interpret that meaning. In this case, this ball is indeed grom, which is a substance created by the dreugh, an underwater race in the north and western parts of the continent. For one year of their life, they undergo karvinasim when they walk upon the land. Following that, they return to the water and meff, or devour the skin and organs they needed for land-dwelling. Then they vomit it up into little balls like this. Grom. Dreugh vomit."

The students looked at the ball a little queasily. Sotha Sil always loved this lesson.

\cnote{4 Sun's Dawn, 2920: The Imperial City, Cyrodiil}
"Spies," muttered the Emperor, sitting in his bath, staring at a lump on his foot. "All around me, traitors and spies."

His mistress Rijja washed his back, her legs wrapped around his waist. She knew after all these many years when to be sensual and when to be sexual. When he was in a mood like this, it was best to be calmly, soothingly, seductively sensual. And not to say a word unless he asked her a direct question.

Which he did: "What do you think when a fellow steps on his Imperial Majesty's foot and says 'I'm sorry, Your Imperial Majesty'? Don't you think 'Pardon me, Your Imperial Majesty' is more appropriate? 'I'm sorry,' well that almost sounds like the bastard Argonian was sorry I am his Imperial Majesty. That he hopes we lose the war with Morrowind, that's what it sounds like."

"What would make you feel better?" asked Rijja. "Would you like him flogged? He is only, as you say, the Battlechief of Soulrest. It would teach him to mind where he's stepping."

"My father would have flogged him. My grandfather would have had him killed," the Emperor grumbled. "But I don't mind if they all step on my feet, provided they respect me. And don't plot against me."

"You must trust someone."

"Only you," smiled the Emperor, turning slightly to give Rijja a kiss. "And my son Juilek, I suppose, though I wish he were a little more cautious."

"And your council, and the Potentate?" asked Rijja.

"A pack of spies and a snake," laughed the Emperor, kissing his mistress again. As they began to make love, he whispered, "As long as you're true, I can handle the world."

\cnote{13 Sun's Dawn, 2920: Mournhold, Morrowind}
Turala stood at the black, bejeweled city gates. A wind howled around her, but she felt nothing.

The Duke had been furious upon hearing his favorite mistress was pregnant and cast her from his sight. She tried again and again to see him, but his guards turned her away. Finally, she returned to her family and told them the truth. If only she had lied and told them she did not know who the father was. A soldier, a wandering adventurer, anyone. But she told them that the father was the Duke, a member of the House Indoril. And they did what she knew they would have to do, as proud members of the House Redoran.

Upon her hand was burned the sign of Expulsion her weeping father had branded on her. But the Duke's cruelty hurt her far more. She looked out the gate and into the wide winter plains. Twisted, sleeping trees and skies without birds. No one in Morrowind would take her in now. She must go far away.

With slow, sad steps, she began her journey.

\cnote{16 Sun's Dawn, 2920: Senchal, Anequina (modern day Elsweyr)}
"What troubles you?" asked Queen Hasaama, noticing her husband's sour mood. At the end of most Lovers' Days he was in an excellent mood, dancing in the ballroom with all the guests, but tonight he retired early. When she found him, he was curled in the bed, frowning.

"That blasted bard's tale about Polydor and Eloisa put me in a rotten state," he growled. "Why did he have to be so depressing?"

"But isn't that the truth of the tale, my dear? Weren't they doomed because of the cruel nature of the world?"

"It doesn't matter what the truth is, he did a rotten job of telling a rotten tale, and I'm not going to let him do it anymore," King Dro'Zel sprang from the bed. His eyes were rheumy with tears. "Where did they say he was from again?"

"I believe Gilverdale in easternmost Valenwood," said the Queen, shaken. "My husband, what are you going to do?"

Dro'Zel was out of the room in a single spring, bounding up the stairs to his tower. If Queen Hasaama knew what her husband was going to do, she did not try to stop him. He had been erratic of late, prone to fits and even occasional seizures. But she never suspected the depths of his madness, and his loathing for the bard and his tale of the wickedness and perversity found in mortal man.

\cnote{19 Sun's Dawn, 2920: Gilverdale, Valenwood}
"Listen to me again," said the old carpenter. "If cell three holds worthless brass, then cell two holds the gold key. If cell one holds the gold key, then cell three hold worthless brass. If cell two holds worthless brass, then cell one holds the gold key."

"I understand," said the lady. "You told me. And so cell one holds the gold key, right?"

"No," said the carpenter. "Let me start from the top."

"Mama?" said the little boy, pulling on his mother's sleeve.

"Just one moment, dear, mother's talking," she said, concentrating on the riddle. "You said 'cell three holds the golden key if cell two holds worthless brass,' right?"

"No," said the carpenter patiently. "Cell three holds worthless brass, if cell two --"

"Mama!" cried the boy. His mother finally looked.

A bright red mist was pouring over the town in a wave, engulfing building after building in its wake. Striding before was a red-skinned giant. The Daedra Molag Bal. He was smiling.

\cnote{29 Sun's Dawn, 2920: Gilverdale, Valenwood}
Almalexia stopped her steed in the vast moor of mud to let him drink from the river. He refused to, even seemed repelled by the water. It struck her as odd: they had been making excellent time from Mournhold, and surely he must be thirsty. She dismounted and joined her retinue.

"Where are we now?" she asked.

One of her ladies pulled out a map. "I thought we were approaching a town called Gilverdale."

Almalexia closed her eyes and opened them again quickly. The vision was too much to bear. As her followers watched, she picked up a piece of brick and a fragment of bone, and clutched them to her heart.

"We must continue on to Artaeum," she said quietly.

The Year continues in First Seed.