\cchapter{Fool's Ebony}{Frincheps}

Dramatis Personae:
\small
\begin{description}
\item Prologue
\item The Adventurer, A Dark Elf Rascal
\item Komon, A Priest of Akatosh
\item Lheban, Another Priest of Akatosh
\item Epilogue
\item Stete, A Priest of Julianos
\item Raic, Another Priest of Julianos
\item Shub, A Mage
\item Shub, A Different Mage of the Same Name
\item Nephron, A Somewhat Sleazy Merchant
\item 5 Armorers
\item Ortho
\item Crunn, Husband of Millie
\item A Lusty Contessa
\item Millie, Innkeep and Philosopher
\item Gurnsey, Bovine Wench
\item Assorted Wenches and Cads of the Taverns
\item Soldiers
\item Dwarves
\item Giants
\end{description}

\csection[Act]{1}

\dropcap{C}{oncerning} \textit{Priests and Nackles As related at length by two Priests of Akatosh to the Adventurer, who at the time was not having an adventure, and had nothing better to do. In which some (probably unwanted) light is shed upon the Priesthood and its members, and upon an old peasant myth of some significance, especially common in High Rock. And in which the mysterious Fools' Ebony appears, that strange material that could bring either drastic cultural change for the many, or just great profit for a few, or death for a bunch, or have no result whatsoever.}

\textit{Daggerfall and Environs in the Doldrums of the 3rd Era}

\textit{Early in the month of Frostfall. The Dead Daedra Inn. Enter Prologue}

PROLOGUE: Our poor players will try and remember their lines and not trip over our meager set. I beg you, the audience, not to heckle, badger, or throw rotten foodstuffs. You will only make this short play last longer. The Guild of Playwrites, Actors, and Dramatists wish any of you who are sensitive or allergic to rambling dialogue, wooden acting, incomprehensible exposition, or unsatisfying endings that leave one confused and unhappy to exit the theatre immediately. Your gold will, alas, not be refunded. As a saving grace, this series of vignettes contains gratuitous references to all pleasures of the flesh. You may enjoy it. Ah, here comes our hero, the roguish Dark Elf called the Adventurer. It is time for Prologue to trip merrily away.

\textit{Exit Prologue}

\textit{Enter the ADVENTURER}

ADVENTURER: What an odd conversation I just heard between those two mages. It is best not to speak of such matters next to privy hedges.

\textit{Enter 2 Priests of Akatosh (LHEBAN, KOMON)}

LHEBAN: Mind if we join you, fellow? ... Good, need some company ourselves. I am named Lheban, my fellow priest here is Komon. We both serve Akatosh, all in our own ways, of course ...

ADVENTURER: Make yourselves at home, it's not my bench. But I thought that priests ... didn't go to ... er ... places like this, Inns. I mean ... unless on duty?

LHEBAN: Oh, we're not on duty. Got to regenerate our internal vital energies, so we can go on blessing and curing ...

KOMON: We often come here, hike up our robes, kick up our heels, as it were. Fill up with some bottled energy ...

\textit{(Komon snickers)}

LHEBAN: Looking for those in need of comfort and blessing, of course ...

KOMON: Oh, yes, Oh yes ... like that young girl outside the other evening ...

\textit{(Lheban kicks Komon)}

KOMON: ... and anyway our High Priest told us to get lost...

LHEBAN: He means told us to get some air. We've been having visions, you see...

KOMON: Yes, sort of weird, really ... and we hadn't even been taking any of that ...

\textit{(Lheban kicks Komon)}

LHEBAN: Both of us been having the same visions - real odd.

ADVENTURER: Do tell, I'm not going anywhere in a hurry.

LHEBAN: Well, we've both been hearing sort of ... words ... for a start. Like 'Sir Nich' or 'Sain Nack' ...

ADVENTURER: You said 'Nick' or 'Nack'? Just a minute ... let me have a swig from your bottle, Brother ... Ah! That's better - high-class stuff you fellows drink! Yes, I recall - some story or old legend about an elf, name of Nuckle, I think -- from Morrowind?

LHEBAN: You know, maybe you're on to something there -- there is a old legend around these parts, comes from deep in High Rock I think ... hmmmm ... Nackles, that's it!

ADVENTURER: Nackles, eh! Seems that several Dark Elves use that name ... particularly the ... more peculiar ones...

KOMON: Yes, I guess that the bad ones are into all that weapons magicka stuff ... very nasty fellows ...

LHEBAN: (to Komon) Komon! This fellow's got pointy ears and red eyes ...

KOMON: Pardon me, friend ... it's sort of dark, and I didn't ... uh ...

ADVENTURER: Oh, that's fine. These are strange times. You know, live and let live -- or die -- as the case may be. Now ... suppose you tell me about this Nackles myth? Here, let me help you with that bottle ... Ah! Thanks.

LHEBAN: Er ... sure, if you want to put it that way ... Here, have another swig! Sure, we've got the time, and I recall it clearly now.

KOMON: Yes, we've a couple hours 'til that little blonde shows up at her lamp...

\textit{(Lheban kicks Komon)}

LHEBAN: (to Komon) Quiet! Remember, we had to tell the High Priest her address, so she won't be around for a while!

(to all) Very well, here's the story, best as I can recall it. This is a tale the peasants up in High Rock tell their kids to scare them into being good for a while, I guess. They tell it, let me see ... either on Tales and Tallows, or is it Witches' Festival? -- just before the kids are sent out to the barn or pigsty to sleep.

KOMON: Nasty cruel peasants! But then, I'd send them all out to the midden ...

LHEBAN: Really, Komon! Remember, those poor souls need our compassion and blessing, we are their salvation!

KOMON: Now who's in Old High Mucky-Mucks' study?

LHEBAN: Er ... anyway. It goes a bit like this. If the kids have been real good during the year -- filched enough in the market, mucked out the stables every day, not gone playing with goblins, left the sheep alone, and so on. If they have been real good, they've nothing to worry about. But if they haven't been real good then there is this nasty, horrid Dark Elf spirit called Nackles. Doesn't look like your typical Dark Elf -- thinner, taller. Pasty white face, long as your arm. Walks like his knees and elbows bend the wrong way. Snickers like when you drag your fingernails across slate. Wears a tight black suit (not Khajiit, more like a formal suit with buttons) but too tight and small. He visits the bad girls and --

KOMON: Why are you talking about Old High Mucky again, Lheban?

\textit{(Komon hiccoughs) (Lheban kicks Komon)}

LHEBAN: You really must excuse Komon here: overwork, you know. Too many curings and conversions ... Anyway, Old Nasty Nackles is supposed to wander under our Tamriel, in dirty deep dark dwarven tunnels. Everywhere under the lands, if you can believe that! Rides in a rusty squeaky old mine cart, on old mine tracks ...

ADVENTURER: I saw some of those in Fang Lair once, down in Hammerfell a long long while ago ...

KOMON: (to Lheban) What the Sheogorath was he doing in Fang Lair!?

LHEBAN: (to Komon) Hush! If he's who I think he is, you do not want to know! (to all) Um, yes. Well, Nackles gets pulled all around these deep tunnels by goblins -- not your usual dirty yellow ones, but nasty black things. Anyway, they pull Nackles round and through these dark tunnels, and then, late at night, he stops below each and every bad child's hovel or house or castle - makes no difference. Then he slides up the drainage pipes ...

KOMON: Creeps up cracks ... crawls through holes ...

LHEBAN: Oozes up oubliettes ...

KOMON: Climbs giggling up garderobes ...

LHEBAN: Right into the kid's place! Then, if the kid's only been sort of bad, Nackles will just mess things up in general, so the kid gets blamed. Make greasy dirty marks everywhere (more than usual, anyway), break some things, steal some things, so on and so forth. Maybe take the sugar sweets, leave some lumps of fools' ebony instead ...

ADVENTURER: Fools' Ebony - what's that? Heard mention of that, oh, a few hours ago ... Some Mages ...

LHEBAN: You did now? Interesting ... Very ... Well, lets talk of that in a bit ... just let me finish this Nackles thing. Where was I -- Oh yes ... Now, if the little brat has been real bad -- then all the little brat's toys get taken. The copper dagger, the wooden sword, the little whip, and so on. All the usual favorite kids things.

KOMON: Whips? I like those.

\textit{(Komon hiccoughs) (Lheban kicks Komon)}

LHEBAN: Now if that little brat has been very, very bad then Nackles grabs the brat. Pops him or her in his dirty great sack. Hauls the sack off down the holes and cracks, down to his rusty old mine cart! And away they go!

KOMON: Hope he leaves some bad little girls behind.

\textit{(Lheban kicks Komon)}

LHEBAN: Er ... so we can save them, of course, friend ... Well. Sometimes, so I've heard tell, the brat never comes back. No great loss, I guess, peasants just breed another.

KOMON: Know 'bout that, I do, I do ...

\textit{(Lheban pinches Komon's nose)}

LHEBAN: But, as the story goes round here anyway, often the brat is just put to work, digging out lumps of Fools' Ebony, shoveling dirt, bagging it. Extending the tunnels of the Nackles. After a while, Brat is pushed back up to where it came from. Seems that Brat might think it's spent a year down there, but only a day has passed up top ... Brat comes back real thin and dirty though, covered in black mess ... You know, come to think of it -- on the day past Witches' Festival, I've often seen some little brats, scrawny, real dirty black mess on them, looking terrified, too. Parents drag them into Temples to get blessed and cured, if they have the gold. By the Beard of Sheogorath, the wailing and noise! Enough to drive a priest to ... er ... well, never mind ... that's our problem ...

KOMON: Nah ... it's a problem with our suppliers, I tell you ...

\textit{(Lheben throws Komon through a screen)}

LHEBAN: Anyway, that's the short of it, this Nackles legend up around here. I recall now, it's widespread all over Tamriel ... and knowing the place, probably more than a grain of truth in the tale, much, much more ...

ADVENTURER: So, I guess some of the ... er, darker Dark Elves sort of identify with this Nackles. Take on the persona, so to say ...

LHEBAN: Yeah, that sort of sums it up, I guess ... though we don't see those types hauling off brats in sacks, now do we?

KOMON: Nah, that's wot we does, girly brats anyway, isn't it not?

\textit{(Komon hiccoughs) (Lheban breaks a bottle over Komon's head)}

\textit{(Komon falls unconscious)}

ADVENTURER: Thats a very interesting tale, gentlemen. Say, let me repay you with another bottle -- what's that you're drinking? Ah, thought so - Innkeep! More holy wine for these holy men!

LHEBAN: A blessing on you for that kind gesture, friend.

ADVENTURER: I thank you, I sure could use one or three ... Anyway, this 'Fools' Ebony', I've heard mutters and murmurs about that of late -- mostly eavesdropping ... pardon me ... listening ... to Mages and the like. What's with this stuff? Here, have another swig ... good!

LHEBAN: Well, we're not supposed to tell outsiders ... but then, you seem to know something already. And if you have been hearing Mage gossip ... Why, maybe we can do some business. Profit all round! Well ... for the Akatosh Chantry, of course, and your fee, good Sir.

ADVENTURER: More and more interesting -- tell on, I pray you.

\textit{(Komon staggers to feet) (Komon hiccoughs)}

KOMON: Time for me to go convert that little lamppost girl ... no, no, no - not last nights one, but the blonde ...

\textit{(Exit Komon) (Female squeals from offstage)}

LHEBAN: Friend, you'll have to excuse Komon. He's a bit ... you know strange ... Got these ...

ADVENTURER: Oh, that's all right, we've all got our own...

\textit{(Exeunt Lheben and the Adventurer) (Enter EPILOGUE)}

EPILOGUE: Our apologies for the quality of this drama so far. If those of you still present will wait for a few minutes while our bard plays "Silence Implies Consent," we will change the set for the next act, Part the Twoth. Please don't forget to tip your wench. Do you believe there's such a thing as Fools' Ebony? Maybe we'll find out in Part the Twoth. Or maybe not.

\textit{(Flourish) (Exit Epilogue)}

\textit{End of Part the Oneth, Being Mostly Concerned with The Legend of Nackles.}