\chapter*{\headerfont Lord Vivec's Sword-Meeting with Cyrus the Restless}

\dropcap{G}{ather}, sit. Drink to Papa and tell Morwha you’re sorry for what you’ve done wickedly this day save for the customary curses allowed towards our enemies, the fair skins and the green skins and the sataks that roll in the dirt with no skins at all. Most importantly, listen, here’s another tale of Sura of the Bend’r Mark, the Maverick-Sword of the Crowns, called Cyrus in the tongue of the Septims.

Mind you, this story isn’t necessarily true, for no tale of Cyrus the Restless is true in its entire, and yet that has never really mattered. Indeed, you’ll come to see that sometimes stories that aren’t necessarily true can still sometimes win the day.Now these were the days after the Tiber War had come to Hammerfell with its banners and phalanxes and its skill at sea, which came to nothing except for treaty, for no empire of men may rule the sons and daughters of Yokuda. We are superior in every way to all the other races of this world and they know it. The Hoon Ding guides us; all others can Make Way.That is, unless those others do not come from this world at all, and are instead spirits of the Far Shores or the In-Between-- against these gods and demons and mad shapers, all men can be humbled. But is that not the point? Different rules apply to gods and demons and mad shapers. Of these, I talk now of the demon-king of the Star-Wounded East, Vehk and Vehk the Circle Talker, anon Ansu-Gurleht in Yoku, the God of Makes Us Women, whose powers are so terrible that even Redguards must bend to His will, though foreign and thus benighted; if He comes, we run until we can’t. There is no shame in this; we are allowed to run from disaster; witness the loss of Yokuda, where our running was blessed by windy Tava who filled our sails with escape. And Ansu-Gurleht is surely disaster.But there was one of us that didn’t run from Him. Cyrus never ran from anything except for the killing of Hakan, and that was so long ago that it doesn’t really count.

\sectionline

The captain had them restow the Carrick at Herne and once more at Jabbur before moving into the waters of the Abecean. The map to Old Yokuda came from the Lame Cat of Wayrest, like always. There was no agenda save the looting of a particular temple; at least that’s all Cyrus would let on about; but some of the raga of the crew were eager just to see the homeland of their forebears. True stories had come that Yokus still lived among the stark remains; that some did not or could not flee when cataclysm came; others that the orichalc isles were a place of censure handed down from the no-totambu. And everyone knew the self-exiled ansu still lived there who did not witness the ho no shira, or the capture of Volen, or the Make Way of Diagna, and so were left to the sword-singing of their histories behind the Curtain of Run.

The captain set their course south-southwest, slanting across the spring trades into the Sea of Pearls to the puzzlement of the crew. Some of the older among them muttered darkly of the guardians of the western approaches, but nothing was sighted during the long weeks of smooth sailing. At twilight on the 12th day out of Jabbur, Coyle, long-learned in the navigations, took sighting of the non-constellation of Sep and abruptly Cyrus changed course to the north. The old hands explained to the new that they had crossed the Line that day and it was now safe to bear up for Old Yokuda. The winds, which had been fair, now turned into the north and gusted with growing ferocity, often threatening to lay the Carrick on her beam if not for the skill of her crew. After weary days of this, they finally made landfall on the 17th day.

\parabreak

\textit{Coyle, stay with the ship and start the trim. If the locals come, you speak the language. Anchor’s dropped right, sir, and Borden’s already got his raiders filed. Good, we move soon, night’s falling. Take Haekele with you, Captain, he speaks Yoku, too, and reads it if he don’t lie. Noted. We go to the spot S’rathra marked and nowhere else. Sen nung ni-Bateki tro ki-lodo. What? It’s a prayer, Cap’n, to the God of We Like Our Bodies Just Like They Are. Fair enough. Ach, heathens help us. No offense, sir. It’s a just tradition, Captain. Aye, these isles tro zhang-ga let. Shut up and get your kit.}

\parabreak

By torchlight they made for the interior, badlands all of it and worse. Sharp drops would appear along the stone pathways or the high wet walls, all of it jagged from the mythic wars against the Aldmer and still gleaming with eldritch-foam. Cyrus was reminded of the geographies of Morrowind and Masser, though the wind smelled sweeter here. He choked out an order for rope to get them across the Tendu Shreds that was coming sooner than his crew had thought possible. Their captain had memorized the map, since he knew they would approach the temple under cover of night. He passed some licorice to Fornower, the youngest of those present.

The temple was nestled in the Shreds, unlit save for the foam that gave it an underwater glow, and Borden said a small word to Tsun to keep his cool. "All in a day’s," he said.

Broken and worked, the temple held a glimmering language above its entrance. Haekele of Alik’r moved closer. Cyrus raised a hand.

"It’s Daedric," the Captain said. "I can read that."

At the mention of the Lords of Misrule, the crew of the Carrick were set to a heavier unease.

"’The Virtue of the Little Reward’," Cyrus read aloud. Then he frowned at the writing. "Well, I’m glad I could work that out for us." And then he guided them in.

\parabreak

\textit{The statue looks dunmer. Shh. But that makes no sense, sir, they ain’t much fer sailin. Just find the jewel if we can, this place looks pretty bare. Doesn’t look ransacked, though. Cap, look there, that’s probably why. It’s a gate. That’d keep anyone out. Hand the torch. Move up, Gar, what’s its aura read? Look on the walls. Shh, what’s it read? We’re fine, sir, it can’t open, it’s ‘fire-locked. The Emperor’s good for something then. I’m serious, look on the walls. Those kings, they aren’t dunmer, they’re raga. No, they’re women, see. All of them is with child. Look. For. The. Jewel. Oh no, Captain, this is a temple to Ansu-Gurleht! Who the hell is 
that?}

\parabreak

They found that the jewel was in parts, three by three, one for each etching of the pregnant Yoku kings of the Temple of Ansu-Gurleht the Seed-Bringer. With careful knives they pried them out and by Gar’s small enchantments they formed them into one, the Glass Opal of the Nogru.

"AI PADHOME ALTADOON!"

Borden took out his sword. "That came from the gate, sir."

Cyrus looked at the arching columns of the temple center, long cracked by age and with traces of wheel-carvings. "Yeah," he said.

The gate spoke again. "Drop the bauble, rude shapes, I am bathing in pico-filament moltings that come in six beautiful colors the Aurbis hasn’t even revealed yet, and I really do not want to make this journey. You should also know that I’m very dangerous."

"Move," Cyrus said, and the crew of the Carrick ran.

\parabreak

The route back to the beach was quicker, as they jumped what they had had to climb before, and the discretion of their movements they put aside as fast as the sun could rise. Some of them thought foolishly that daylight might keep their pursuer away, as he was undoubtedly a dark thing perhaps not given to new mornings, but Cyrus yelled at them to run. Better than Gar, he could feel when strange things were moving against him.

They ran across the sand towards the boats and their shipmates knew enough to start the casting off. "Pull her up," Coyle yelled, "They’ve caused trouble!"

The wind fell around them, and then a shadow, and then a bright half-star behind. Cyrus turned to see, and maybe he was smiling because he liked this kind of thing, and from the rock edges Vivec appeared, making the men gasp. No one had expected a dunmer on these shores, nor one so oddly arraigned.

"I’ll deal with him," Cyrus said, and Borden waited a second to receive the Opal, but his captain wouldn’t toss it. "No hard feelings, but I don’t really trust you, Borden. Go!"

Vivec was half-golden and half-blue and all of him glowing, and he was armored lightly, and carried a small shield and a curved sword at his side. His head was bald except for flame, and he smiled with evil.

"Stay back, dunmer," Cyrus said, eyeing his head. "I can’t have you burning my boat."

Vivec walked closer. Afterward, men aboard the Carrick said they could smell a fortune in bug-musk.

"You don’t know me," Cyrus said, "so I’ll tell you once--"

And then Vivec spoke, and the winds of the beach died down when he did for he was the Lord of the Middle Air and they were indentured to him. He said to Cyrus, "Oh, I know you, raga. I know that you forced an armistice with the Cyrodiil, which I have done, though by other designs. I know also that even after speaking well and rendering your people free again, you remained a thief at heart in the days after; these days, in fact, which I have also done and still do, again in my own way."

And here the god of the East smiled a bit too lovingly. "And let it be said here that you have no idea how much I absolutely adore thieves."

Cyrus for his part was circling around Vivec, making a trough in the sand, some place in which to move easier when the fight came. His men watched from the deck of his ship. Some had brought bows and arrows, but Coyle moved them away with a hand.

Vivec lifted his legs to float in the lotus position, his head to the side with the smile vanished and replaced by a dole of remorse without mocking. "I know how you die," he said, "and the trouble your soul will have reaching the far shores of your taken stars because of things you did to the discredit of the Hist, and how their long roots run even into the void tendril-feeling for your final entrance. I know how you think now, at this moment, that there are no paths except for the drowned lamp, or the wrongheaded romance of saberplay in a landscape of long regret, taking whosoever will ride with you through the still-sought salvation spread across each water lash, wandering your heart to find some purchase beyond the admonishment of the moons; flagellant without end."

Vivec’s eyes went to the Glass Opal cradled in the pirate’s arm and frowned. He looked at Cyrus, dourly, saying, "I know all of this about you, Sura, and more, and it grants us a kinship despite your crime against me, and so it pains me yet that I think you really know nothing of me and my mastery. Or do you? I am the city that walks, the wise and benevolent eye of the ALMSIVI, and it has been such for uncounted red generations, each adding its mark then and forevermore, the worship of the construction-everlasting, Architect Amen. What could you possibly know of me?"

Cyrus had never changed his expression. "I know you’re talking," he said, "because I see your mouth moving and I hear words." And with that he dropped the stolen jewel and drew his saber.

Vivec then drew his own sword, slow yet perfectly, whipping around to angle it to his left. "Hmm," he said. "Yes, fine, I think."

The two moved in closer, Vivec gently floating so that the bonemold of his right armor faced Cyrus. "Death despite kinship has ever been our way," he said, "I know too how it feels to murder the husband of my sister."

At this last, Cyrus was finally angered, understanding now that this champion of Morrowind was truly able to read his mind, which he had vouchsafed from even the closest to him, and this is why he hated all spirits of aether. It read on his face, which thereafter hardened. "Tell me, demon," he asked, "since you know how I die: is it this fight?"

"No."

"Good to hear," Cyrus said, and attacked.

Vivec spun in midair cross-legged, never moving his swordarm from its initial position. He spoke, "Fa-Nuit-Hen," the name of an old master, and Cyrus fell from eight wounds that appeared without mortal notice. Coyle could not stop the men from firing then.

Bleeding into the sand, Cyrus could see Vivec above him with no sword in his hand at all but instead the stolen jewel of the Nogru, and an array of seventeen arrows fanned around his firehead aspect peacock-style, caught by demon magic. Cyrus could not get up and Vivec spoke, "And I know of your late father, the playwright, and though some of its local color is lost on me, I am fond of his work. That is why I have let you live. I adore poetry, too."

The Lord of the Middle Air vanished, and Coyle sent boats to the shore with medicines and their sugarcat surgeon, who had taken passage sometime after Herne.

\parabreak

\textit{You cannot go back. We must cast off. Captain, take us into this no more. A demon’s mercy only comes once. Quit staring at your sword, sir. Yokuda was lost for a reason. Let’s just go. Why not just send the crows an invitation in lights. Please, captain, let us cast off. I mean, we brought some from Akavir, they’d see ‘em surely. You can’t even move. We’ve heard the stories of Ansu-Gurleht, the raga among us at least; the gods say it’s okay to hightail it. Great work, Haekele, those were right magicsome words. Go get the skooma-pipe and tell him it’s a scalpel day again. Please let us just leave this haunted place. Cut up suits you, cap, pregnant don’t; listen to the men. You cannot go back. With all due respect, sir, you’re dumber than I look.}

\parabreak

Now in his time asleep or under the knife, Cyrus didn’t know that Coyle had set out with a few other sailors, all of them Yoku-speakers. Knowing his captain too well, the young man went to find the fabled ansus, the sword saints of orichalc. Coyle kept no hope that he should be so lucky—the ansus kept their citadels hidden, and komodos protected them under color of rock. In the village of Bu-tabar, which topples down the eltheric breaks like survivor pearls, Coyle learned that the ansus were seeking to hold counsel with Cyrus anyway. He met them on the Samarand Road and they were three as in tradition- a young one, an aged one, and a daughter of neither. They had a train of followers to carry their sword collections and their memories-in-stone. Seeing the crew of the Carrick, the daughter ansu reenacted the Falls of Diag-leeki with twin falchions, playing every part; since she did so in under a minute, it meant hello-don’t-be-useless.

"I should warn you," Coyle told the elder in the old tongue, "Captain Cyrus doesn’t know any Yoku at all."

\parabreak

Coyle brought the elder ansu and the young one to meet Cyrus in the surgeon’s hold; the daughter stayed abovedecks with their train, to protect them and to warn the sailors with displays. The young ansu greeted the captain in Yoku, which Coyle translated for him. Cyrus merely nodded and lifted up, bandaged across his eight wounds, and the sugarcat hissed at the needless movement. Startled by the sound, the young ansu took a Walled Consequence stance. Coyle brushed the surgeon away quickly and finally the elder spoke.

"We know of you, Surahoon," he said in the captain’s own speech, "And motions were multiplied the other morning on this beach, which is sword tremor, which we can feel, and that can only mean you encountered the Ansu-Gurleht."

Cyrus shrugged his shoulders. "If you mean the dunmer wizard, yeah, he was here. I almost had him. And no funny talk."

The young ansu dropped his stance. "He cannot be beaten," he told Cyrus, "We know every sword move created in history, and none of them would avail you, even if you learned them from our memories-in-stone, which we would be obliged to lend you if you asked. You killed the white king in the Hammerfell, after all."

"What?" Cyrus said. "The Emperor? I didn’t kill him."

"Of course you did; you were the Hoon Ding."

"No I didn’t and no I wasn’t."

The young ansu refused to listen. Behind his stone-feather mask he smiled in admiration. "You disarmed him, even, and would not kill him until he showed another knife. That is ra gada honor. We do not fight the unprotected. Your stories have come—"

"That wasn’t the Emperor," Cyrus said. "That was just—"

"Of course it was. That is why the Hammerfell stands. You were the Hoon Ding. In any case, the Ansu-Gurleht cannot be beaten. He was gifted by the Barons of Move Like This, who record sword moves from the future, as well."

"Right," Cyrus sighed. "Please remove yourself from my ship."

Coyle went to his captain. "Sura," he whispered, "nogo tur—"

"Not an option, Coyle." Cyrus turned to the ansu. "Now go. The fight with the dunmer is my own. I appreciate the advice but you’re—"

"Will you not take our memories-in-stone, at least?"

"Why?" Cyrus said. He tried to keep his temper. "You’ve already said that they’re useless against him."

The elder lifted a hand. "The Hoon Ding could—"

"Look, I am not –"

"The Hoon Ding could read the stones and show you what we say is true—that no move exists that will get past the Ansu-Gurleht."

The elder looked through the floor of the hold. "At least not now," he said, "The Hoon Ding would show you that he will not manifest in you in this fight."

"Good," Cyrus said. "He’d get in my way." And with that he lay back down. The elder looked up from the floor, and set his head to the side.

"Surahoon," he said, "We are the ansu, the greatest warriors that live in men. Our swords sent the Left-Handers into the oceans, whose empire was four times the size of the white king. When we fight, our swords can kill the laws of nature itself. Yokuda is as you see it because our hira-dirg swords can cut the atomos, the uncuttable, and we did. We are the ansu, and we tell you now that you cannot beat the Ansu-Gurleht. How do you think he came by that name? Who do you think was our finest student?"

Coyle stepped back. He shook his head. Cyrus for his part just shrugged again.

"Then your finest student is a painted hussy," he said, "And a foreign one at that. Great work. He’s quick enough, sure, but I’ve cut down his kind before."

Cyrus lifted his head for a second, looking to Coyle. "He talks a lot, too, and seems to like that. Is there a God of Talk to Death here, too?"

"Yes," Coyle said, "But Ansu-Gurleht is the—"

"Of course there is," Cyrus said, "I can work with that, then. Hopefully he’ll be holding his sword as he’s yakking away so my Redguard honor will remain intact when I run my saber into his mouth mid-sentence."

The young ansu spoke with the elder’s voice. "We are telling the Hoon Ding to run."

Cyrus smiled and closed his eyes.

"Where’s the money in that?"

\parabreak

\textit{Bellguard down, over, hold. The Bone Shaver. Strike at 80 grams, any degree but this one. The Ephemeral Feint. Breathe in and then forget the breath; you cannot replace it until he is down, to fight as if dead: second principle of pneumansu. The Vectoring Cygnet. Arm out, knee down, coal on the teeth to hide your smile. The Pankratosword, but this is forbidden. Arc the bones that otherwise cannot bend. The Threat of Mirrors. Using the Math Athlete, you could occur several places during a single duel, illustrious and sure. Paint fake eyes all over your face and then hide your real ones among them; the opponent can no longer read where you look. The Premeditated Modesty. The Fingers-Knife serves as five, protecting your cardinal points and your central theory; five thrusts, spaced microseconds apart, like tapping the desk bored, waiting for morning bread.}

\parabreak

Cyrus woke in the surgeon’s hold, dark lapping sounds coming through wood. The cat was still up going through bottles and washing linens. Seeing his captain wince, he nodded and said, “Bad moons in a big dream.”

“Tell me about it.”

“Before you even ask,” the cat said, “G’latha slipped no moonsugar in your last resting-water.”

Cyrus moved to a sitting position, wincing slightly from some pain in his lower back. “Then why?” he said.

“The sword-walkers left a memory-stone under your pillow.”

Cyrus moved the pillow and saw it: a dark rock weathered smooth with age, encrusted with traces of glistening curves. He glared at G’latha. “And you knew about this?”

“Yes,” the cat said. “G’latha found it amusing. Also, G’latha thought that maybe its magic would seep into your brainpan and teach you vaba maaszi lhajiito, do-sura.”

Cyrus got on his feet and grunted some admonishment. He found his sword near the cutter tools and hefted it, looking at its grip and wondering why it felt a bit off. Had the dunmer wizard knocked its balance out? Had he even set aside the saber in that whirling motion he’d made? Cyrus couldn’t remember. He remembered only how ridiculously fast it had all been.

“That book also teaches ahzirr traajijazeri,” he said to G’latha.

“Did you ever notice we hid that lesson at the bottom? Speaking of lessons, it seems like do-sura learned a few.”

“Yeah, there was a move called—oww.” Cyrus had attempted a swing but something across his skin tore sharply. “I was going to make a joke in there somewhere about skinning —oww. I reopened something.”

“G’latha was serious. Did you not notice what you did?”

Cyrus put his saber down. He was feeling dizzy. The sugarcat lent a hand.

“Here, sit back down for the needle,” G’latha said. Pawing his captain’s stitches lightly, his eyes flashed over to the sword.

“Do-sura fights right handed,” he said, voice low, “Yet you just held your saber with your left.”

\parabreak

As morning rose, Cyrus scanned the beach again from topside. Some men were moving about nervously. Thorpe, the scrub, was close by, swabbing the decks, whistling some song from Sutch or thereabouts.

“Thugs mustn’t have ties to anything except ink and the glimmer of gold.”

Thorpe looked up from his brush. “What was that, sir?”

“Nothing,” Cyrus said, realizing he’d been heard. “Something an old villain of mine said to me once.”

“Ring a truth it’s got,” Thorpe said, going back to his work, “And if ye don’t mind me sayin’ so, there’s plenty of glimmer back homewards.” Thorpe started nodding at his own advice. “Scuttle’s that the Reachers have taken to the water to fight Old Mary,” he continued, “and those Bret’n buttertubs couldn’t outrun the Carrick on their best day.”

“That’s just it, Thorpe,” Cryus sighed. "The Reachers don’t stand a chance against the Dominion. But they’re trying. Same news you heard says the Thalmor sent a plague into Camlorn. It’s a matter of…" And Cyrus let the sentiment trail off. He had taken an interest in the sunrise.

"What’s that, sir?” Thorpe said, eyeing him. "Reconciling vengeance and honor again?”

"Yoku has forty-eight different versions of honor,” Cyrus said, "And they all trace their roots to ugak-ta, which means, more or less, "I’m pretty mad”. So that might be a yeah.”

Fornower walked by, fetching a saw. "Woke up speaking Yoku, cap?” he said.

"Strange night,” Cyrus said.

Thorpe let Fornower pass until he spoke again. "Ain’t a soul on this boat would think less of ye fer avoidin’ that whip-evil Velothi domino, Cy.”

"I know.”

Cyrus palmed the top of his saber-hilt, thinking. He frowned as he felt the same imbalance in the blade, even in its sheathe. He sighed again and looked over to Thorpe.

"Actually, there would be one that would think less of me.”

\parabreak

So it was that Cyrus the Restless came to a decision, which was ever preordained by his make, and let us not say that he did so easily, for that would be a lie, but still and alas: he was to meet the Ansu-Gurleht in single combat for satisfaction, if only the demon-king might be troubled again to show up. Cyrus had thought of ways to ensure this.

"Bring me maps and histories of the dunmer," he told the learned of his men, "And bring any here that know of the Ansu-Gurleht's aspect beyond local legend."

Borden showed with documents, for he was a reaver long, and had been to Morrowind before and after Armistice. He knew nothing much, however, of the dark elves beyond their ability to fight at raid time. "He's not a Redoran, I can say that," Borden said, scrolling out papered territories and pointing, "He was little armored, and those devils go weighted."

It was Gar that knew most of their dunmer politic, but he could not reveal much that wasn't tangled in old tale. "He may be the one named Vivec, because he mentioned ALMSIVI, which is some kind of god guiser cabal in their belief, with great hold on their nation. This "Vivec" is the only one of that group that frolics about outside their borders so much; one might say he's their public face or something."

Cyrus snorted.

"I'm telling you just what I know, Captain. The Pocketguide says they never come out at all, the ALMSIVI. Better texts link him to the Daedroth Mephala."

"Oh," Borden added, "Them Redorans also speak less distinguished than the dandyman you...well, I would say you fought, but that'd be generous in description. No offense, as always."

"Mephala?" Cyrus said. "Which Lord is that?"

Gar flipped through a book, looking for something to back up what he was already going to say. "The Webspinner in day parlance; otherwise it gets hazy. Ties to the Morag Tong, ties to sordid other cults in the provinces, running gamut from drugs to dibbledark to, hell, fashion trends even. Weird one, this Mephala."

"Weird is relative," Cyrus quoted.

A crowd had gathered; men with rope or hidden eagerness or barely covered discomfort. Their captain was stubborn and famous for it. No matter the disposition, each Carricker needed to be a part. Coyle was among them, about to say something until Hiddleman spoke up.

"Prizewise, what yer talkin past all this raga noise is we get the Nogru Opal back, yeah?"

The men of the Carrick shifted then, eyeing their captain and the dumb marine from Anvil. Coyle took the stead. He said, "Yes, Hiddle. No action Sura decides goes tobr'a way. Settle in and give scandal no glance."

Cyrus took the heed, but his nature couldn't stop his color despite it. "Listen, all of you. Runners were sent inland, and they're coming back with trade. The locals have never seen fireworks, and we stole plenty from Ko. Orichalc's en route in abundance, which will gather drake like even the oldest here have never witnessed. S'rathra's shiny is only going to make us richer, and I aim to get it back from that dunmer princess without a doubt."

Cyrus looked at the men to see the measure of their waver. "Without a doubt, you hear," he said. "Because no power lays me low like that on any ground, foreign or however-you-like. I’m Redguard. He insulted me, and let me live. In plain, don't trouble on my motive and your payday’ll come tonight."

He looked to Borden. "Bring me my armor." He turned to Gar. "Find me the worst thing I might say to this Vivec." He turned to Coyle. "Load the cutter. I go within the hour."

\parabreak

With no need for secrecy this time, Cyrus went a more direct route to the temple of the Ansu-Gurleht. Near a crossroads that split on the westernmost fall of the Tendus there was a wayside shrine to Morwha, with a sprawling statue inside of a fashion he had seen before in his own lands. He left coin there in a wax-hollowed candle holder on the goddess' knees and felt a sudden relief in the wound of his lower back. He nodded to the ugly, fat face of the Yoku mother spirits and gave thanks and moved on.

The day became hot, for he was unused to wearing chainmail and helm and greaves, and heavy padding underneath all of it. He'd done so at times, battled in armor, but even at war in M'kai or the Rim he went without it and relied on speed instead, because his fighting style did. There would be no speed against the demon-king of the dunmer, Cyrus had learned, only force and hard defense. His mind was heady with the sword-thought of the ansus. It was all of it breath and geometry of edge and lessons that were lost on him because he had not the physical training for them.

He wondered about the Barons of Move Like This and the impossibility of their designs; how they were an advantage of Vivec's that few warriors could compensate for. He wondered how that any stone-knowledge he'd absorbed had already been dealt with by a countermove that did not exist yet; and that the Ansu-Gurleht had no doubt trained in it. Finally, he wondered of various ways to just plain cheat.

"There is nothing you can say to insult this Tribune, Captain," Gar had said. "He is divine and goes all back all eras. He has endured slights from every power-- of the dark, and of the starkeepers, and of the lesser royalty of Man-- and delivered them back with laughter. His main joy is foul effrontery, one could say, and will gain you nothing."

Cyrus told Gar he wasn't helping.

"Actually, I'm trying," Gar said. "More than insulting him with something unlikely to earn anything but amused contempt, I'm afraid that he would counter with something worse. He has a godsmind, Captain. He knows he could unshake you, and you cannot suffer that to happen in your duel with him."

\parabreak

It had somehow rained only inland the past day. The earth near the Shredstart was soft and unlovely, broken by stream. Cyrus stopped and looked down a wide gulf, thinking in Yoku, netu anselim, which was the small version of the thought "turn back". Below, salmon leapt upwards along a low river in their unparsible function.

Another Yoku phrase almost rose, but Cyrus smothered it, saying only, "Yes, that's just about right." The salmon and their desperate natal stretch was an answer he agreed to.

He set the unease of the stone-knowledge into his stomach, removing it for a moment to become again what he remembered himself to be and not what magic had wrought, the ancestor idioms breaking across a bank inside him.

He cursed the ansus for their interference. No one should be made to swallow a language whole. He looked to the fish again and shrugged, finally moving on. Yes, he thought, it's only a fool that picks which dooms he subscribes to if he has more than one that breach his sky...but, well, there you go.

"I'm coming for you, woman," he said, quickening.

\parabreak

And when he saw that the entrance to the temple had been sealed by a seamless plane of ebony, Cyrus wondered if he had come for nothing. Then all thoughts of honor went away and he wondered instead if the Carrick had anything aboard that might cut godsblood, because he would be stinking rich.

Then the right half of the ebony plane turned to gold, and both halves moved like fluid in the air, a vision of trapped rain raining along a hectic compass; and Vivec formed from this mercurial swirling and floated there, finally smiling. He had not changed in his arraignment, Cyrus noted. And he notices that I have. Great. Hush.

Vivec spoke. "So the ansus I sent didn't ward you off, little cut? I willed them to, for I have mentioned my liking of thieves, and thus I sought to help you, raga doon, Cyrus the Restless."

"Tung den uta-no-mongo, Ansu-Gurleht."

Vivec cocked his head, which lit aflame at the wondering. "Well, something stuck, at least. And I give help whether you wish it or not. What I will not give is my toy of the westernmost west, the Opal of the Nogru, whose chieftains I lady-made in long ages before your kind came to the Tamri-el."

"We'll see about that," Cyrus said.

"Tenacity," Vivec said, drawing his sword swiftly, "I adore that, too. I am the circumvention of the Black Hands of excess, and now have the rule of it, and it was not bought without tenacity. May I name you as wife after this is over?"

And with that the Ansu-Gurleht angled his sword down in the position he had done at the beach, point down, blade edge forward, 60 degrees rise-wise from the ground. He floated lotus fashion and stopped smiling. His eyes were watching the future, Cyrus knew, so think of a way around all of this, damn it.

Cyrus drew his saber, holding it left-handed in the Tempest Attrition.

Vivec was impressed and said so. "That stance allows you a riposte," he said. Then he smiled again. "If you're quick enough."

Cyrus moved to Havoc Canal and Vivec nodded. "Better. Your build favors that. But what if I've enchanted my skin to harden at each instance?"

Zero-sum Interrupt.

"Can you even do that, silly raga?"

"Test me."

"No," Vivec said, "For I would prefer to imagine that you could. It would mean that I would bleed prior to the cut itself, bypassing all healing; moreso, and this is what pleases me utmost... it would mean that I have indeed helped you."

Cyrus had his moment then, where he had maneuvered the duel from the start: pride. Pride was the Ansu-Gurleht's weakness. "Milord," Cyrus said, "Forgive me my little grace, but I deny your help again. Indeed, I turn my back to it, and betray it, and malign it. That should be familiar to you."

Vivec paused. "It is," he said.

"As should this," Cyrus said, moving to the Pankratosword.

Vivec paused longer. And then he laughed loud. "You would not!"

"I say again, test me."

"You would destroy the home of your ancestors even more? And in the fashion that they had done, which is now forbidden in your hands?"

Cyrus didn't move from his stance.

"All for an Opal which you could never thereafter spend?"

"You would be finished," Cyrus said. "And I would be rid of the one who shamed me. So why wouldn't I? You shouldn't have sent your saints to me, Ansu-Gurleht. I am not much like them at all."

"Cut the atomos and you die, too."

"What, the Barons of Move Like This didn't teach you a countermove to this?" Cyrus said. "Oh, wait, they wouldn't have. The Pankratosword is stricken from the record."

Vivec laughed again. He was delighted. He said, "The things they said about you were true, Redguard, whether you wish to believe it or not."

Vivec bowed his head.

"I make way," he said. "I drop my sword."

And the demon-king did, though it hovered a few inches from the ground.

Vivec continued. "And thus we come to stalemate, for I am unarmed and you cannot--"

"No," Cyrus said, moving his sword to Vivec's throat, "You overestimate my Redguard honor, boy. Different rules apply to gods and demons and mad shapers. And you, Ansu-Gurleht, you're all three."

"I should kiss you," Vivec smiled.

"The jewel'd do better," Cyrus said, "Thank you very much. Summon it, send it to my ship, and we're done."

"This is what you ransom for your shame," Vivec said, chiding, the Opal of Nogru floating out from within the temple. It flew off south and east. "You're easily bought."

Cyrus turned to go, but Vivec stopped him, saying,"And I had so wanted to see the Pankratosword, and so believed you ired enough to bring its ruin on us both. What stopped you, Surahoon?"

"Old advice," Cyrus said. And he pointed his saber to Vivec's face and its two colors. "And you're wearing it, sure enough."

At the apex of the temple's upper reaches, Cyrus turned back, calling to Vivec. "Ink and gold, milord," he said. "And the general mystery of dangerous men, you see. Well met, Tribune, and farewell."

He left the Ansu-Gurleht forever.

And that's how the story goes, with a duel that was not a duel, and with a story that wasn't true. Cyrus, you see, never knew how to actually use the Pankratosword, only how to hold it at threat. After all, its use was forbidden and thus held in no stone at all after the fall of Old Yokuda.

When he returned to his ship, the Carrickers cheered and asked loudly over one another the same questions: how the fight went with the Ansu-Gurleht, and how their captain had become the victor. Cyrus smiled then, and wouldn't answer, moving them to hustle and trim, for they were casting off.

It was Fornower who got the truth of it, handing Cyrus half a bit of licorice. "For your mighty deeds, now plus one: some bitten-on licorice of the Bay" he said. "But, c'mon, Cap, and tell just me, then. How'd you beat him?"

"Yeah, well," Cyrus said. He chewed and swallowed. "I cheated."